### 2024-06-13 output/manuscript.tex
Updated

\documentclass[12pt]{article}

\usepackage{latex/ediArticle}  %from ediArticle.sty

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\LARGE \bfseries \vspace{-1cm} Just Another Latex Template \\
    \LARGE \normalfont Research Proposal}
\author{\Large \textit{Edifofon Akpan}}
\date{ }

\begin{document}

% Set font size to 13pt
\fontsize{13pt}{15pt}\selectfont

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Lorem ipsum dolor sit amet

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
Lorem ipsum dolor sit amet

\hspace{1em}

\begin{table}[H]
    \centering 
    \caption{The facts about lorem ipsum generator from online sources.}
    \begin{tabular}{|C{2cm}|L{10cm}|C{2.5cm}|}
        \hline
        \PlainInput{tables/char-included-studies}
    \end{tabular}
    \label{MyTable1}
    \caption*{\footnotesize \textit{Notes:} Donec sit amet viverra justo. Duis malesuada velit}
\end{table}

\section{Results}
Lorem ipsum dolor sit amet (\autoref{MyTable1}).

\hspace{1em}

\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{output/figures/dose-price-income.png}
    \caption{The facts about lorem ipsum generator from online sources.}
    \label{fig:enter-label}
    \caption*{\footnotesize \textit{Notes:} Lorem ipsum dolor sit amet}
\end{figure}

\section{Discussion}
Lorem ipsum dolor sit amet.\parencite{aljunid_economic_2014}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% %%%
%%% Bibliography %%%

\clearpage
\newrefcontext[sorting=nyt]
\printbibliography

\end{document}