### 2024-06-30 output/tables/tab_likely_switch.tex
Updated

Characteristic & Threshold & Most likely switch\\
\hline
Type of comparison & Ochalek (2018) high & PCV7 vs NoVax (35.0\%, p=0.665)\\
\hline
Type of comparison & Ochalek (2018) low & PCV10 vs NoVax (48.4\%, p=0.514)\\
\hline
Type of comparison & Woods (2016) high & PCV10 vs NoVax (35.9\%, p=0.562)\\
\hline
Type of comparison & Woods (2016) low & PCV10 vs NoVax (78.1\%, p=0.043)\\
\hline
Type of comparison & P-Rivière (2023) high & PCV13 vs NoVax (26.5\%, p=0.961)\\
\hline
Type of comparison & P-Rivière (2023) low & PCV10 vs NoVax (53.1\%, p=0.764)\\
\hline
Income group & Ochalek (2018) high & Low (57.1\%, p=0.031)\\
\hline
Income group & Ochalek (2018) low & Low (57.1\%, p=0.122)\\
\hline
Income group & Woods (2016) high & Low (57.1\%, p=0.021)\\
\hline
Income group & Woods (2016) low & Lower-middle (85.7\%, p=0.056)\\
\hline
Income group & P-Rivière (2023) high & Low (57.1\%, p=0.003)\\
\hline
Income group & P-Rivière (2023) low & Lower-middle (60.7\%, p=0.240)\\
\hline
Herd effects & Ochalek (2018) high & Excluded (32.7\%, p=0.544)\\
\hline
Herd effects & Ochalek (2018) low & Excluded (43.0\%, p=0.706)\\
\hline
Herd effects & Woods (2016) high & Excluded (31.8\%, p=1.000)\\
\hline
Herd effects & Woods (2016) low & Excluded (72.9\%, p=0.155)\\
\hline
Herd effects & P-Rivière (2023) high & Excluded (24.3\%, p=1.000)\\
\hline
Herd effects & P-Rivière (2023) low & Excluded (51.4\%, p=0.350)\\
\hline
Serotype replacement & Ochalek (2018) high & Included (52.9\%, p=0.050)\\
\hline
Serotype replacement & Ochalek (2018) low & Included (52.9\%, p=0.433)\\
\hline
Serotype replacement & Woods (2016) high & Included (41.2\%, p=0.409)\\
\hline
Serotype replacement & Woods (2016) low & Included (82.4\%, p=0.271)\\
\hline
Serotype replacement & P-Rivière (2023) high & Included (41.2\%, p=0.126)\\
\hline
Serotype replacement & P-Rivière (2023) low & Excluded (48.8\%, p=1.000)\\
\hline
Study perspective & Ochalek (2018) high & Healthcare/payer (31.0\%, p=1.000)\\
\hline
Study perspective & Ochalek (2018) low & Healthcare/payer (44.0\%, p=0.611)\\
\hline
Study perspective & Woods (2016) high & Healthcare/payer (33.3\%, p=0.595)\\
\hline
Study perspective & Woods (2016) low & Societal (71.0\%, p=0.720)\\
\hline
Study perspective & P-Rivière (2023) high & Societal (25.8\%, p=0.698)\\
\hline
Study perspective & P-Rivière (2023) low & Healthcare/payer (51.2\%, p=0.506)\\
\hline